Temperature and wave exposure as drivers of morphology in two kelp species

### 2019-07-16 paper_morph_new.Rmd
shallow morphology and significance tests added

---
title: Temperature and wave exposure as drivers of morphology in two kelp species
  along the coast of South Africa
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
geometry: margin=1in
output:
  word_document:
    fig_caption: yes
  pdf_document: default
papersize: A4
tables: yes
fontsize: 10pt
---

@@

# Results
## Temperature variables
```{r software packages, message=TRUE, warning=TRUE, include=FALSE, echo=FALSE}
library(tidyverse)
library(ggpubr)
library(gridExtra)
```

@@

```{r Temperature variables plot, echo=FALSE, warning=FALSE, fig.cap="Temperature variables at the collection sites around the Cape Peninsula. Temperature variables include minimum represented by blue dots.", fig.height=9, fig.width=7}

@@

```{r wave data prep, message=FALSE, warning=FALSE, include=FALSE}

@@

```{r wave variable processing, warning=FALSE, include=FALSE, echo = FALSE}

@@

```{r wave variable plot, warming = FALSE, echo=FALSE, message = FALSE, include= TRUE, fig.cap='Wave variables at each collection site around the Western Cape coast. Mean variables are represented by dots and standard deviation by whiskers. Plot "A" represents annual significant wave height.', fig.height=8, fig.width=7}

@@

```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE}
lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>% 
   filter(site!= "Rocky Bank")

@@

pos_temp <- c("Cool temperate", "Warm temperate")

@@

  stat_compare_means() + 

@@

lam_df$wave_exposure <- factor(lam_df$wave_exposure, levels=c("Very sheltered", "Very exposed", "Moderately exposed","Fully sheltered","Extremely sheltered"), labels=c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed"))
pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed")

LW1 <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Lamina weight (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level"))

@@

  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +

@@

```{r laminaria temp plots, echo=FALSE, message=FALSE, warning=FALSE}
ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom", label = "p.format", ncol = 3, nrow = 3)
```

```{r laminaria wave plots, echo=FALSE, message=FALSE, warning=FALSE}
ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom")
```

@@

```{r ecklonia  processing, message=FALSE, warning=FALSE, include=FALSE}
library(ggplot2)

## Preparing data
# Ecklonia
eck_data <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
 filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
 filter(site!= "De Hoop")  
# group_by(wave_exposure, frond_mass) %>% 
# summarise(mean = mean(frond_mass:total_length, na.rm = TRUE)) %>% 
# ungroup()

eck_deep <- eck_data %>% 
  filter(population == "Deep")
eck_shall <- eck_data %>% 
  filter(population == "Shallow")

@@

# custom positions for wave exposure categories and sites
pos_exposure <- c("Fully sheltered", "Extremely sheltered", "Very sheltered","Moderately exposed", "Exposed", "Very exposed")

frond_mass_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_mass, fill = wave_exposure)) +

@@

  stat_compare_means() + 

@@

primary_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +

@@

primary_width_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +

@@

frond_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +

@@

stipe_mass_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +

@@

stipe_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +

@@

stipe_circ_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_circ, fill = wave_exposure)) +

@@

tufts_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +

@@

epi_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +

@@

total_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +

@@

```{r ecklonia plots, echo=FALSE, message=FALSE, warning=FALSE}

ggarrange(frond_mass_plot,primary_length_plot,primary_width_plot,frond_length_plot,stipe_mass_plot,stipe_length_plot,stipe_circ_plot, tufts_plot, epi_length_plot,total_length_plot, labels = "AUTO", common.legend = TRUE, legend = "bottom")

@@

>>>>

---
title: Temperature and wave exposure as drivers of morphology in two kelp species
  along the coast of South Africa
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
geometry: margin=1in
output:
  pdf_document: 
    fig_crop: no
  word_document:
    fig_caption: yes
papersize: A4
tables: yes
fontsize: 10pt
---

@@

# Results
## Temperature variables

```{r software packages, message=TRUE, warning=FALSE, include=FALSE, echo=FALSE}
library(tidyverse)
library(ggpubr)
library(gridExtra)
```

@@

```{r Temperature variables plot, echo=FALSE, warning=FALSE, fig.cap="Temperature variables at the collection sites around the Cape Peninsula. Temperature variables include minimum represented by blue dots.", fig.height=8, fig.width=8}

@@

```{r wave data prep, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}

@@

```{r wave variable processing, warning=FALSE, include=FALSE, echo = FALSE, message=FALSE}

@@

```{r wave variable plot, warming = FALSE, echo=FALSE, message = FALSE, include= TRUE, fig.cap='Wave variables at each collection site around the Western Cape coast. Mean variables are represented by dots and standard deviation by whiskers.', fig.height=8, fig.width=8}

@@

```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>% 
   filter(site!= "Rocky Bank")

@@

pos_temp <- c("Cool temperate", "Warm temperate")#custom temperature categories position on x-axis
#temperature and lamanaria  

@@

  stat_compare_means(method = "wilcox.test") + 

@@

#wave exposure and lamanaria
lam_df$wave_exposure <- factor(lam_df$wave_exposure, levels=c("Very sheltered", "Very exposed", "Moderately exposed","Fully sheltered","Extremely sheltered"), labels=c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed"))
pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Moderately exposed", "Very exposed")

LW1 <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
  stat_compare_means(method = "kruskal.test") + 
  ylab("Lamina weight (g)") + 

@@

  stat_compare_means(method = "kruskal.test") + 

@@

```{r laminaria temp plots, echo=FALSE, message=FALSE, warning=FALSE, fig.height=9, fig.width=9, fig.cap='Boxplot of morphological characteristics for *L. pallida* categorised by temperature regime.'}
ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom", label = "p.format", ncol = 3, nrow = 3)
```

```{r laminaria wave plots, echo=FALSE, message=FALSE, warning=FALSE, fig.height=9, fig.width=9, fig.cap ='Boxplot of morphological characteristics for *L. pallida* categorised by wave exposure.'}
ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c('A', 'B', 'C','D','E','F','G','H',"I"), common.legend = TRUE, legend = "bottom")
```

@@

```{r deep ecklonia  processing, message=FALSE, warning=FALSE, include=FALSE}
library(ggplot2)
## Preparing data
# Ecklonia
eck_data <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
 filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
 filter(site!= "De Hoop")  
# group_by(wave_exposure, frond_mass) %>% 
# summarise(mean = mean(frond_mass:total_length, na.rm = TRUE)) %>% 
# ungroup()
eck_deep <- eck_data %>% 
  filter(population == "Deep")
eck_shall <- eck_data %>% 
  filter(population == "Shallow")
pos_temp <- c("Cool temperate", "Warm temperate") #custom temperature categories position on x-axis
#temperature plots made seperately for each morphological characteristic
ET1 <- ggplot(data = eck_deep, aes(x = temp_regime, y = frond_mass, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
ET2 <- ggplot(data = eck_deep, aes(x = temp_regime, y = primary_length, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
ET3 <- ggplot(data = eck_deep, aes(x = temp_regime, y = primary_width, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
ET4 <- ggplot(data = eck_deep, aes(x = temp_regime, y = frond_length, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
ET5 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_mass, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
ET6 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_length, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
ET7 <- ggplot(data = eck_deep, aes(x = temp_regime, y = stipe_circ, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  ylab("Stipe circ (cm)") +
ET8 <- ggplot(data = eck_deep, aes(x = temp_regime, y = tufts, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  ylab("No. of tufts") +
ET9 <- ggplot(data = eck_deep, aes(x = temp_regime, y = epi_length, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
ET10 <- ggplot(data = eck_deep, aes(x = temp_regime, y = total_length, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 

@@

# custom positions for wave exposure categories and sites
pos_exposure <- c("Fully sheltered", "Extremely sheltered", "Very sheltered","Moderately exposed", "Exposed", "Very exposed")
#morphology and wave exposure
EW1 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_mass, fill = wave_exposure)) +

@@

  stat_compare_means(method = "kruskal.test") + 

@@

EW2 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +

@@

EW3 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +

@@

EW4 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +

@@

EW5 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +

@@

EW6 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +

@@

EW7 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_circ, fill = wave_exposure)) +

@@

EW8 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +

@@

EW9 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +

@@

EW10 <- ggplot(data = eck_deep, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +

@@

```{r deep ecklonia temp plots, echo=FALSE, message=TRUE, warning=FALSE, fig.height=9, fig.width=9, fig.cap ='Boxplot of morphological characteristics for deep population *E. maxima* categorised by temperature regime.'}
ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
```

```{r deep ecklonia wave plots, echo=FALSE, message=FALSE, warning=FALSE, fig.height=9, fig.width=9, fig.cap = 'Boxplot of morphological characteristics for deep population *E. maxima* categorised by wave exposure.'}
ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8, EW9, EW10,labels = "AUTO", common.legend = TRUE, legend = "bottom")
```

```{r shallow ecklonia temp processing, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
EST1 <- ggplot(data = eck_shall, aes(x = temp_regime, y = frond_mass, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
EST2 <- ggplot(data = eck_shall, aes(x = temp_regime, y = primary_length, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
EST3 <- ggplot(data = eck_shall, aes(x = temp_regime, y = primary_width, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
EST4 <- ggplot(data = eck_shall, aes(x = temp_regime, y = frond_length, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
EST5 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_mass, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
EST6 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_length, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
EST7 <- ggplot(data = eck_shall, aes(x = temp_regime, y = stipe_circ, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  ylab("Stipe circ (cm)") +
EST8 <- ggplot(data = eck_shall, aes(x = temp_regime, y = tufts, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  ylab("No. of tufts") +
EST9 <- ggplot(data = eck_shall, aes(x = temp_regime, y = epi_length, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
EST10 <- ggplot(data = eck_shall, aes(x = temp_regime, y = total_length, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 

@@

```{r shallow ecklonia wave processing, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
eck_shall$wave_exposure <- factor(eck_shall$wave_exposure, levels=c("Fully sheltered", "Extremely sheltered", "Sheltered", "Exposed", "Very exposed"), labels=c("Fully sheltered", "Extremely sheltered", "Sheltered", "Exposed", "Very exposed"))
# custom positions for wave exposure categories and sites
pos_exposure_shall <- c("Fully sheltered", "Extremely sheltered", "Sheltered", "Exposed", "Very exposed")
#morphology and wave exposure
ESW1 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = frond_mass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_shall) + 
  stat_compare_means(method = "kruskal.test") + 
ESW2 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_shall) + 
  stat_compare_means(method = "kruskal.test") + 
ESW3 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_shall) + 
  stat_compare_means(method = "kruskal.test") + 
ESW4 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_shall) + 
  stat_compare_means(method = "kruskal.test") + 
ESW5 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_shall) + 
  stat_compare_means(method = "kruskal.test") + 
ESW6 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_shall) + 
  stat_compare_means(method = "kruskal.test") + 
ESW7 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = stipe_circ, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_shall) + 
  stat_compare_means(method = "kruskal.test") + 
ESW8 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_shall) + 
  stat_compare_means(method = "kruskal.test") + 
ESW9 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_shall) + 
  stat_compare_means(method = "kruskal.test") + 
ESW10 <- ggplot(data = eck_shall, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_shall) + 
  stat_compare_means(method = "kruskal.test") + 

```{r shallow ecklonia temp plots, echo=FALSE, message=TRUE, warning=FALSE, fig.height=9, fig.width=9, fig.cap ='Boxplot of morphological characteristics for shallow population *E. maxima* categorised by temperature regime.'}
ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
```

```{r shallow eck wave plot, message=FALSE,warning=FALSE,include=TRUE, fig.height=9, fig.width=9 ,fig.cap = 'Boxplot of morphological characteristics for shallow population *E. maxima* categorised by wave exposure.'}
ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, ESW10,labels = "AUTO", common.legend = TRUE, legend = "bottom")
```
